MLH1 Data Report

### 2019-09-11 doc/Main_Report.rmd
fixed buggs and removing un-needed sections

## Description of Report

- Show basics of the current data file

- Display plots and figures of distributions of MLH1 counts to assess initial overall patterns.

-  Report basic statistics of results and to help keep track of the number of images for each category and mouse.

@@

```{r quant.repeatability, echo=FALSE}

@@

there is a positive correlation with MLH1 number and quality
To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.

@@

Some papers have documented age effects for male genome-wide recombination rate (cite).

@@

```{r draw.rand.cells, echo=FALSE}
ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
```

@@

table(chrm.class.DF$category)
str(chrm.class.DF)

@@

)
str(this_thing2)

#this first one is the 'first' MLH1 count
oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
summary(oopp2)
#none of the classes are significant

#mean is the new calculated-- mean 0CO is significant and intercept
oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
summary(oopp)
#not sure if these are inter-correlated
#I don't know why 0COs and the intercept are significant..
#0CO and 1CO are significant when only chrm classes are included (and intercept)

@@

chrm.class.DF$biv.sum.check

@@

SC.skel$rand_name <- as.character(SC.skel$rand_name)
original_DF$Random.Name <- as.character(original_DF$Random.Name)
str(SC.skel)
str(original_DF)

@@

str(mlh1.skel.og)
mlh1.skel.og$skel_size  <- as.numeric(mlh1.skel.og$skel_size)
mlh1.skel.og$bin_size <- as.numeric(mlh1.skel.og$bin_size)
mlh1.skel.og$nMLH1.foci <- as.numeric(mlh1.skel.og$nMLH1.foci)

@@

str(new_shiny)
str(MSM_BivData)

@@

ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
  geom_boxplot(aes(alpha=.3))

@@

biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
theme(legend.position="none")+facet_wrap(~strain)

>>>>

@@

```{r quant.repeatability, echo=FALSE, include=FALSE}

@@

There is a positive correlation with MLH1 number and quality
To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.

```{r quality.effect, echo=FALSE}

setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
#compare all quality 1 cells across strains
#check that the means (or directions of differences is maintained)

Q12_MLH1_DF <- MLH1_data[MLH1_data$quality <= 2,]

#ddply
Q12_MLH1_means <- ddply(.data=Q12_MLH1_DF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

tot.DF_MLH1_means <- ddply(.data=MLH1_data, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

mergy <- merge(Q12_MLH1_means, tot.DF_MLH1_means, by='category', all = F)
colnames(mergy) <- c("category","Q12ncells","Q12mice","Q12_MLH1_means","Q12MLH1var",
                     "Totncells","Totnmice","TotMLH1mean","TotMLH1var")

mergy$percent.diff <- (mergy$Q12_MLH1_means / mergy$TotMLH1mean)

Q12toTotal <- ggplot(mergy, aes(y=Q12_MLH1_means, x=TotMLH1mean))+geom_point(aes(size = Q12ncells) )+geom_abline(slope = 1, intercept = 0)+ggtitle("Quality 1 cells MLH1 mean to total MLH1 mean")+ylim(c(21,32))+xlim(c(21,32))

Q12toTotal

#check that the directions are the same
#make tables for the 
#is the amount of 'polymorphism' the same

Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "female",], 
                 .(category,subsp),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci)
)

tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "female",], 
                 .(category,subsp),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci)
)

kable(Q12_MLH1_means_males, caption = "High quality cells")

kable(tot.DF_MLH1_means_males, caption = "Total Cell")

All the quality 1 cells produce a higher mean. The biggest difference in means is HMI male, LEW female and SPIC female means have ~10% differences.

When quality 1 to 2 is compared, the strains were closer to the line, but the means are still higher in the higher quality cells.
with 1-2 most all of the difference are below 10% (except SPRET female).

MOLF males have quite a big difference.

There seem to be the similar amount of polymorphism in the Dom and more conserved in Dom.

Ranges for male means across subspecies. Ranges of means (not ranges of all).

(try turning this into a table...)
Dom            female Dom
old,   new      old     new
22-24  23-25    24-27  25-29

Musc           female Musc
23-31   25-31   25-27   26-29
22-25   22-27    25       25
(ASTmale has 27 mean, outlier)

Outgroup       female outgroup
25-26   26-27    27    29-31

I think the ranges across the groups

@@

Some papers have documented age effects for male genome-wide recombination rate (cite, one of the Cole papers).

@@

```{r draw.rand.cells, echo=FALSE}
#ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
#ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
```

@@

#table(chrm.class.DF$category)
#str(chrm.class.DF)

@@

)
#str(this_thing2)

#this first one is the 'first' MLH1 count
oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
#summary(oopp2)
#none of the classes are significant

#mean is the new calculated-- mean 0CO is significant and intercept
oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
#summary(oopp)
#not sure if these are inter-correlated
#I don't know why 0COs and the intercept are significant..
#0CO and 1CO are significant when only chrm classes are included (and intercept)

@@

#chrm.class.DF$biv.sum.check
```

@@

SC.skel$rand_name <- as.character(SC.skel$rand_name)
original_DF$Random.Name <- as.character(original_DF$Random.Name)
#str(SC.skel)
#str(original_DF)

@@

#str(mlh1.skel.og)
mlh1.skel.og$skel_size  <- as.numeric(mlh1.skel.og$skel_size)
mlh1.skel.og$bin_size <- as.numeric(mlh1.skel.og$bin_size)
mlh1.skel.og$nMLH1.foci <- as.numeric(mlh1.skel.og$nMLH1.foci)

@@

@@

#ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
#  geom_boxplot(aes(alpha=.3))

@@

biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( ))+
theme(legend.position="none")+facet_wrap(~strain)
